Claret: Using Data Types for Highly Concurrent Distributed Transactions

### 2015-02-17 rmd/paper.Rmd
wordsmithing

    In this work, we demonstrate several ways of leveraging abstract data type (ADT) semantics in databases, focusing primarily on commutativity. Using Retwis, a Twitter clone built for Redis, as a case study, we show that using commutativity can reduce transaction abort rates for high-contention, update-heavy workloads that arise in real social networks. We conclude that ADTs are a good abstraction for database records, providing a safe and expressive programming model with ample opportunities for optimization, which will make databases more safe and scalable.
---
```{r setup, include=FALSE}
opts_chunk$set(dev='pdf', echo=F, message=F, warning=F, error=F, fig.width=3.6, fig.height=3)
```

@@

**Probabilistic data types** are a class of data types with *probabilistic* guarantees about their semantics, trading off some accuracy for better performance or storage. These include *bloom filters* [@bloom], *hyperloglog* [@hyperloglog], and *count-min sketch* [@countminsketch]. Hyperloglog, which also appears in Redis [@redis], estimates the size of a set within a fixed error bound. Twitter's streaming analytics system [@summingbird] leverages these to handle their high data volume, and we can expect similar improvements.

**Conflict-free replicated data types (CRDTs)**, which were invented for eventual consistency, can actually be fit into our model as well, as a new kind of data type. This could allow copies to exist in different shards, asynchronously updating each other. By defining the same kind of *merge* function as traditional CRDTs, these replicas could ensure they all converge to the same state. Clients of these might find them more difficult to reason about but may consider making this tradeoff in parts of the application where it otherwise cannot scale.

>>>>

  In this work, we explore several ways of leveraging abstract data type (ADT) semantics in databases, focusing primarily on commutativity. Using a Twitter clone as a case study, we show that using commutativity can reduce transaction abort rates for high-contention, update-heavy workloads that arise in real social networks. We conclude that ADTs are a good abstraction for database records, providing a safe and expressive programming model with ample opportunities for optimization, which will make databases more safe and scalable.
---
```{r setup, include=FALSE}
opts_chunk$set(dev='pdf', echo=F, message=F, warning=F, error=F, fig.width=3.6, fig.height=3)
```

@@

**Probabilistic data types** are a class of data types with *probabilistic* guarantees about their semantics, trading off some accuracy for better performance or storage. These include *bloom filters* [@bloom], *hyperloglog* [@hyperloglog], and *count-min sketch* [@countminsketch]. Hyperloglog, which also appears in Redis [@redis], estimates the size of a set within a fixed error bound. Twitter's streaming analytics system [@summingbird] leverages these to handle their high data volume, and we expect similar benefit.

**Conflict-free replicated data types (CRDTs)**, which were invented for eventual consistency, can actually be fit into our model as well, as a new kind of data type. This could allow copies to exist in different shards, asynchronously updating each other. By defining the same kind of *merge* function as traditional CRDTs, these replicas could ensure they all converge to the same state. Clients may find them more difficult to reason about but might make this tradeoff in parts of the application where it otherwise cannot scale.
